---
title: "Simulation Excercise: Exponential Distribution"
author: "Andrés Hidalgo Vargas"
date: "14/7/2020"
output: pdf_document
---

# Overview

# Simulations

First we set a seed so we can have a constant generation of pseudorandom numbers, making the plots replicable.

```{r}
set.seed(3081997)
```
After the seed is set, we define our variables. The exponential distribution depends on the rate `lambda`, the sample size will be set to 40.
```{r}
lambda <- 0.2
n <- 40
means = NULL
```
The simulation is run to generate 1000 runs of 40 random exponential numbers with lambda 0.2.
The `mean` is taken from each run and stored.
```{r}
for (i in 1:1000){
  means = c(means, mean(rexp(n,lambda)))
}
```

A one sample with 1000 observations is generated to later illustrate the distribution behaviour. 
```{r}
set.seed(3081997)
onesample <- rexp(1000,0.2)
```

# Sample vs theoretical mean

The theoretical mean is defined by 1/lambda. Therefore, the theoretical mean for the simulated data is

```{r}
1/lambda
```

The sample mean is

```{r}
mean(means)
```

4.99 is very close to 5. 


# Sample vs theoretical variance

The theoretical variance is defined by (1/lambda^2)/n. Therefore, the theoretical variance for the simulated data is

```{r}
(1/lambda^2)/n
```

The sample variance is 

```{r}
var(means)
```

0.67 is close to the theoretical variance 0.625

# Distribution

To explain how the distibution appears normal, first we need to see how does the original exponential distribution looks like 

```{r}
library(ggplot2)
ggplot(data.frame(onesample), aes(onesample)) + 
  geom_histogram(aes(y=..density..),
                 binwidth=2,col="black",
                 fill = "lightblue" ,
                 alpha = 0.5) + 
  xlab("X") + 
  ylab("Density") + 
  labs(title = "Exponential distribution of 1000 observations", subtitle = "lambda = 0.2")
```
This distribution looks similar to the [probability density function graph][1]

Now, the distribution of the averages of 40 random exponentials and 1000 simulations will be plotted.
```{r}

ggplot(data.frame(means), aes(x = means)) + 
  geom_histogram(aes(y = ..density..),
                 binwidth=.25,
                 col="black",fill = "lightblue",
                 alpha = 0.5) + 
  geom_segment(aes(x = mean(means), y = 0, xend = mean(means), yend = 0.6), 
               color = "darkred", size=1) + 
  stat_function(fun=dnorm, 
                args = list(mean=5,sd=sd(means))) + 
  geom_text(x = mean(means), label = "\nsample mean", y= 0.01) + 
  xlab("Distribution average") + 
  ylab("Density") + 
  labs(title = "Distribution of the average of 40 random exponentials (1000 simulations)", 
       subtitle = "Normal Distribution (mean = 5 , sd = 0.819)")
```
 As the CLT says, with sufficient random samples, the distribution of the sample means will be approximately normally distributed. 
 
[1]: https://en.wikipedia.org/wiki/Exponential_distribution#/media/File:Exponential_probability_density.svg





